---
title: "ggformula Worksheet"
format: pdf
output:
  pdf_document: default
  html_document: default
authors: Casey Crary and Tyler McCord
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
library(mosaic)
library(ggplot2)
library(RColorBrewer)
library(nasaweather)
library(mosaicData)
```

## Aesthetics (color, size, shape, alpha)

Using the `storms` data frame from the `nasaweather` package, create a scatterplot between `wind` and `pressure`, with color being used to distinguish the `type` of storm.  

```{r}

```

## Axis Labels

Save your above plot, then add appropriate axis labels

```{r}

```

## Facets

The RailTrail data set from the mosaicData package describes the usage of a rail trail in Western Massachusetts. Use these data to answer the following questions. (Hint: For information on how to use faceting and add regression lines, see MDSR, Ch 3.2.2)

1. Create a scatterplot of the number of crossings per day volume against the high temperature that day  
2. Separate your plot into facets by weekday (an indicator of weekend/holiday vs. weekday)  
3. Add regression lines to the two facets  

```{r}
#| warning: false
#| message: false

```

## Geometries (boxplot)

Using the `iris` dataset make a boxplot of `Sepal.Length` across the different species.  What can you conclude?

```{r}

```



