---
title: "Introduction to Regression Modelling"
subtitle: "A Brief Overview of Linear Regression"
author: "Casey Crary and Tyler McCord"
date: 2025-09-10
date-format: iso
format: revealjs
editor: source
---

```{r}
#| label: setup
#| include: false
library(tidyverse)
library(boot)
library(ggplot2)
library(dplyr)
library(mosaicData)
library(ggplot2)
library(mosaic)
library(ggformula)
```

## What is Regression?

-   Regression finds relationships between independent predictors on a continuous numeric scale
-   Reveals more complexity than hypothesis testing
-   ex. What are the factors that predict SAT test scores?

## Our Dataset - SAT

-   SAT data assembled for a statistics education journal article on the link between SAT scores and measures of educational expenditures
-   contains data from 1994-1995

```{r, echo=TRUE}
head(SAT)
```

## Simple Linear Regression

-   We can see how **one** quantitative variable impacts another quantitative variable
-   Find a least squares regression line
    -   Minimize the sum of squared residuals
-   Let's try predicting `sat` (each state's average SAT score) using `expend` (expenditure per pupil in average daily attendance in public elementary and secondary schools, in thousands of US dollars)

```{r, echo=TRUE}
gf_point(data = SAT, sat ~ expend) +
  labs(title = "Scatterplot of State Expenditure versus SAT Scores")
```

## 

```{r, echo=TRUE}
gf_point(data = SAT, sat ~ expend) +
  labs(title = "Scatterplot of State Expenditure versus SAT Scores")
```

## Fitting the SLR Model - 1

```{r, echo = TRUE}
slr <- lm(sat ~ expend, data = SAT)    
msummary(slr)
```

## Fitting the SLR Model - 2

-   Use `lm()` to fit the model

-   `msummary()` to find necessary information

    -   Estimate - Coefficients for each variable, this looks like y=mx + b

    -   Pr (\>\|t\|) indicates the p-value for each predictor, minimize

    -   Residual standard error - every data point is, on average, this far away from the line

    -   Multiple R-Squared - the amount of variation in y that is explained by x, maximize

    -   F-statistic p-value - significance of the entire model, minimize

## SLR Plotted

```{r, echo=TRUE}
gf_point(sat ~ expend, data = SAT) %>%
  gf_lm() +
  labs(title = "Scatterplot of State Expenditure versus SAT Scores") 
```

## Conditions (LINE) - 1

-   Linearity: predictor(s) and response should have a linear relationship

```{r, echo=TRUE}
mplot(slr, which = 1)
```

## Conditions (LINE) - 2

-   Independence: observations should be independent of each other, this has to do with experimental design
-   Normality: We want the residuals to be normally distributed. Use a qqplot or check distribution of residuals.

```{r, echo=TRUE}
mplot(slr, which = 2)
```

## Conditions (LINE) - 3

-   Equal Variance: residuals should have consistent variation

```{r, echo = TRUE}
mplot(slr, which = 1)
```

## SLR Interpretation

-   Fitted equation: $\hat{sat} = -20.89*expend + 1089.29$

    -   For every 1000 dollar increase in expenditure per pupil in public elementary and secondary schools, we predict SAT scores to decrease by 20.89 points

    -   It doesn't make sense to interpret the intercept here.

-   `expend` has a low p-value (0.0064) for `sat`

    -   This p-value is less than alpha=0.05, so therefore `expend` is a significant predictor for `sat`.

    -   Significance of a predictor means that it has a non-zero relationship with the response.

-   Multiple R-Squared = 0.145

    -   This means that 14.5% of the variability in SAT scores can be explained by our model.

```{r, echo=TRUE}
msummary(slr)
```


## Advantages of Linear Regression

-   Easily interperatable

-   Computational time is low

-   Easy to understand

## Disadvantages of Linear Regression

-   Conditions

-   Strongly affected by outliers

-   Real-world relationships are often not 1-to-1

## Sources

-   Thanks to Chelsea Wang for help with the original slides.
-   <https://hbr.org/2015/11/a-refresher-on-regression-analysis>
-   <https://www.geeksforgeeks.org/add-regression-line-to-ggplot2-plot-in-r/>
-   <https://www.statisticssolutions.com/free-resources/directory-of-statistical-analyses/assumptions-of-linear-regression/>
-   <https://online.stat.psu.edu/stat462/node/91/>
-   [https://medium.com/\@satyavishnumolakala/linear-regression-pros-cons-62085314aef0](https://medium.com/@satyavishnumolakala/linear-regression-pros-cons-62085314aef0){.uri}
-   <https://www.geeksforgeeks.org/types-of-regression-techniques/>
-   <https://www.geeksforgeeks.org/understanding-logistic-regression/>
-   <https://www.geeksforgeeks.org/implementation-of-lasso-regression-from-scratch-using-python/>
